---
title: "Bayesian Hierarchical Regression for Ordinal Data"
author: "Peter Hoff"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Bayesian Hierarchical Regression for Ordinal Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides functions to approximate the posterior
distribution of the parameters in the hierarchical
regression  model given by
\[
\begin{align*}
y_{i,j} & = f(z_{i,j})  \\
z_i & =  X_i b_i + e^z_i,   \\
\text{vec}(b_i) & = \text{vec}(b_0)  + V_0^{1/2} e^b_i
\end{align*}
\]
where $f$ is an (unknown) nondecreasing function and

- $z_i\in \mathbb R^{n\times 1}$;
- $X_i\in \mathbb R^{n\times p}$;
- $b_i\in \mathbb R^{p\times 1}$.


for $i=1,\ldots,r$.
The error terms $e^z_i$ and $e^b_i$
consist of independent standard normal entries, and
are also independent
across values of $i$.

The estimation procedure treats $f$ semiparametrically using the
extended rank likelihood. As a result, 
an intercept and scale for the $z_{i,j}$'s are not identifiable.




### load package
```{r}
library(hregression)
```



### generate some parameters and data
```{r}
set.seed(1) 

p<-5 ; n<-50 ; r<-100

X<-rsan(c(r,p,n)) 

b0.0<-rnorm(p)  
V0.0<-rwish(diag(p)) 
ve.0<-1

# latent data
B.0<-Z.0<-NULL 
for(i in 1:r)
{
  b<-b0.0 + t(chol(V0.0))%*%rnorm(p) 
  z<-t(b)%*%X[i,,] + rnorm(n)*sqrt(ve.0) 
  Z.0<-rbind(Z.0,z) 
  B.0<-rbind(B.0,t(b)) 
}


# ordinal data
K<-1+rpois(1,10) 
f<-rnorm(K)*sd(Z.0) 
Y<-matrix( apply(outer( Z.0 , f, ">"),c(1,2),sum) , r, n )
```


### starting values 
```{r} 
Z<-zscores(Y)

B<-NULL 
for(i in 1:r)
{
  B<-rbind(B, lm(Z[i,]~ -1+t(X[i,,]))$coef ) 
}

b0<-apply(B,2,mean) 
```

### MCMC 
```{r,fig.keep='last',results='hide'}
BPS<-B*0 
b0PS<-V0PS<-NULL
KT<-NULL

for(s in 1:1000)
{
  
  V0<-rV0_fc(B,b0) 

  b0<-rb0_fc(B,V0) 

  B<-rB_fc(Z,X,b0,V0)

  Z<-rz_fc(Y,Z,amsprod(X,B)) 

  # output  
  if(s%%10==0) 
  {
    BPS<-BPS+B*10
    b0PS<-rbind(b0PS,c(b0))
    V0PS<-rbind(V0PS,V0[upper.tri(V0,diag=TRUE)]  ) 
    KT<-c(KT,kendalltau(Z.0,Z))

    par(mfrow=c(3,2),mar=c(3,3,1,1),mgp=c(1.75,.75,0))
    plot(B.0,B) ; abline(0,1) 
    plot(B.0,BPS/s) ; abline(0,1) 
    matplot(b0PS,type="l",lty=2,col=1:length(b0)) 
    abline(h=b0.0,col=1:length(b0)) 
    abline(h=apply(b0PS,2,mean),lty=2,col=1:length(b0))
    plot(V0.0[upper.tri(V0,diag=TRUE)], apply(V0PS,2,mean)) ; abline(0,1) 
    matplot(V0PS,type="l",lty=2) 
    plot(KT,type="l") 
    cat(s,mean(Z^2)/mean(Z.0^2),"\n") 
  }

}
```
